---
title: "Color Analysis"
author: "Eric R. Scott"
date: "2022-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load packages and data --------------------------------------------------
library(colorspace)
library(tidyverse)
library(ggrepel)
library(performance)

full_data <- read_rds("data/tweet_data_2022-10-09.rds")
```

# Is the difference between colors a main driver of how close a match is?

In other words, do colors that are more different from eachother have a winner that wins by a greater number of votes?

To understand this, first we need a metric of how different or similar colors are.  I pulled the hex codes for each crayon color from @sbarolo's tweets (assuming they are accurate and official Crayola hex colors).  I then transformed them to the LUV color space, which is relatively "percepturally uniform" meaning that we can use the Euclidean distance between colors in this space as a measure of similarity.

```{r}
#Create matrix column of HSV coordinates for each color
dist_diff <- 
  full_data |>
  mutate(coords = colors |> hex2RGB() |> as("LUV") |> coords()) |> 
  #For each match ...
  group_by(match) |> 
  #Use the matrix to calculate euclidean distance
  summarise(
    distance = as.numeric(dist(coords, method = "euclidean")), #color similarity
    poll_margin = abs(diff(votes)), #how many votes did the winner win by?
    winner = color_name[which.max(votes)], #which color won?
    total_votes = sum(votes), #how many total votes?
    division = unique(division), #what division were they in?
    match_name = glue::glue_collapse(color_name, sep = " vs. ")
  ) |> 
  mutate(margin_prop = poll_margin/total_votes) |> #percentage win margin
  select(match, match_name, winner, division, poll_margin, total_votes, margin_prop, distance)
head(dist_diff)
```

Then we can plot these data:

```{r}
#manual color scale for division
division_color <- 
  c(Blue = "blue", Grayscale = "grey", Green = "green", `Orange/Brown` = "brown", Pink = "pink", Purple = "purple", Red = "red", Yellow = "goldenrod")

ggplot(dist_diff, aes(x = distance, y = margin_prop, color = division)) + 
  geom_point() +
  geom_label_repel(aes(label = match), show.legend = FALSE) +
  # geom_label(aes(label = match)) +
  scale_color_manual(values = division_color) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Are #crayonbracket win margins related to color similarity?",
       subtitle = "labels = match #",
       x = "Color difference (distance in LUV space)",
       y = "% Win (win margin / total votes)") +
  theme_dark() +
  theme(text = element_text(size = 12))
```

It doesn't look like there is any particularly strong pattern, but let's fit a model.

Votes are a count variable, so maybe they're Poisson distributed?  I'm going to go with a Poisson GLM and use an offset of total votes to account for variation in "sampling effort"

```{r}
m <- 
  glm(poll_margin ~ distance,
      family = poisson,
      offset = total_votes,
      data = dist_diff,
      maxit = 100)
plot(m)
performance::check_overdispersion(m)
performance::check_model(m) #there's an NaN in rstandard(m) that causes this to error
```



